---
title: "Details"
author: "Alexis Brunaux"
date: "27/09/2020"
output: html_document
---

The package "alexpackage" allows you to study and summary datas from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. You can doqnload the datas at the url:
https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)

Here below, we will introduce you the functions of this package

#' Convert a CSV file in a dataframe, in order to manage it in R.
#'
#' This function can be used to convert a CSV file in a dataframe, in order 
#' to manage it in R.
#' This function stops and returns a message if the file name does not exist.
#'
#' @param The only parameter of this function is the name of a CSV file which we want to read in R.
#'      (using the \code{filename} argument).
#' @return This function returns a dataframe created from a CSV data file.
#'      This function stops and returns a message if the file name does not exist.
#'
#' @examples
#' \dontrun{fars_read("data/accident_2013.csv.bz2")}
#' \dontrun{fars_read("data/accident_2018.csv.bz2")}
#'
#' @importFrom readr
#'
#' @export 
```{r}
fars_read <- function(filename) {
        if(!file.exists(filename))
                stop("file '", filename, "' does not exist")
        data <- suppressMessages({
                readr::read_csv(filename, progress = FALSE)
        })
        dplyr::tbl_df(data)
}
```

#' Build filename with the year
#' 
#' This filename will be used in the function above fars_read.
#' 
#' @param This function takes only one argument a 4 digits number for the year,
#'      (using the \code{year} argument).
#'
#' @return This function returns simply the name of the CSV file, which we want 
#'      to convert in dataframe with the function above fars_read.
#'
#' @examples
#' \dontrun{make_filename(2013)}
#' \dontrun{make_filename(2018)}
#' 
#' @export 
```{r}
make_filename <- function(year) {
        year <- as.integer(year)
        sprintf("accident_%d.csv.bz2", year)
}
```

#' Create a filename and convert it in dataframe.
#' 
#' This function is a combination of the 2 functions make_filename and fars_read.
#' First it creates the name of a CSV file using the 4 digits of the year argument
#' and using the functionn make_filename..
#' Then it convert the CSV file in a dataframe which we can manage in R, 
#' using the function fars_read.
#' Finally with the functions mutate and select of the dplyr package, a new column
#' with the year and select the columns of the months and the year.
#' If it doesn't exist a filne for the input year, a warning message is returned.
#' 
#' @param This function takes only one argument a vector of several years for the selected years,
#'      (using the \code{years} argument).
#'      
#' @return This function returns a fataframe where the year and the months,
#' which are specified.
#' If it doesn't exist a filne for the input year, a warning message is returned.
#' 
#' @examples
#' \dontrun{fars_read_years(c(2013:2015))}
#' \dontrun{fars_read_years(c(2015:2018))}
#' 
#' @importFrom dplyr
#' 
#' @export
```{r}
fars_read_years <- function(years) {
        lapply(years, function(year) {
                file <- make_filename(year)
                tryCatch({
                        dat <- fars_read(file)
                        dplyr::mutate(dat, year = year) %>% 
                                dplyr::select(MONTH, year)
                }, error = function(e) {
                        warning("invalid year: ", year)
                        return(NULL)
                })
        })
}
```

#' Summarize the number of accidents
#' 
#' This function allows to summarize the number of events (accidents) per months and year.
#' First it reads the CSV file of the selected years and convert it in a dataframe
#' using the function above: fars_read_years.
#' Then it binds the columns and group the lines per month and years using the package deplyr.
#' Finally, it summarize the number of events (accidents) per month and years using the package deplyr.
#' And reorganize the dataframe using the function spread of the package tidyr.
#' 
#' @param This function takes only one argument a vector of several years for the selected years,
#'      (using the \code{years} argument).
#'      
#' @return This function returns a dataframe, which summarizes the number of events
#'      (accidents) per year.
#'      
#' @examples
#' \dontrun{fars_summarize_years(c(2013:2015))}
#' \dontrun{fars_summarize_years(c(2015:2018))}
#'
#' @importFrom dplyr
#' @importFrom tidyr
#'
#' @export
```{r}
fars_summarize_years <- function(years) {
        dat_list <- fars_read_years(years)
        dplyr::bind_rows(dat_list) %>% 
                dplyr::group_by(year, MONTH) %>% 
                dplyr::summarize(n = n()) %>%
                tidyr::spread(year, n)
}
```

#' Build a map where the events (accidents) occurs
#' 
#' This function converts a CSV file in a dataframe and select a specified year and state number, 
#' then using this dataframe builds a map in order to indicate where the accidents occurs and how many accidents occurs in one place.
#' 
#' @param this function take two arguments, first the \code{state.num} argument, which speciies
#'     the state which we are interested.
#' @param the second argument is the \code{year} argument which specifies the year which we are interested.
#' 
#' @return this function returns a map where we can see how many accidents occurs and where they occurs.
#'     If the input state number doesn't exist in the data the function running stops
#'     and the message "invalid STATE number: " is returned
#'     If there is not data for specified state number and year 
#'     the message "no accidents to plot" is returned. 
#'     
#' @examples
#' \dontrun{fars_map_state(c(18,2013))}
#' \dontrun{fars_map_state(c(18,2018))}
#' \dontrun{fars_map_state(c(82:2013))}
#'
#' @importFrom dplyr
#' @importFrom maps
#' @importFrom graphics
#'
#' @export
```{r}
fars_map_state <- function(state.num, year) {
        filename <- make_filename(year)
        data <- fars_read(filename)
        state.num <- as.integer(state.num)
        if(!(state.num %in% unique(data$STATE)))
                stop("invalid STATE number: ", state.num)
        data.sub <- dplyr::filter(data, STATE == state.num)
        if(nrow(data.sub) == 0L) {
                message("no accidents to plot")
                return(invisible(NULL))
        }
        is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
        is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
        with(data.sub, {
                maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
                          xlim = range(LONGITUD, na.rm = TRUE))
                graphics::points(LONGITUD, LATITUDE, pch = 46)
        })
}
```


